R Notebook

### 2021-06-09 paper-calculations.Rmd
some cleaning

```{r}
selectivv <- readRDS("data/selectvv-data-list.rds")
str(selectivv, 2)
```

BAEL data 
```{r}
bael <- read_excel("data/bael.xlsx", sheet = 1)  %>%
  na.omit() %>%
  unite(period, year:yearhalf, sep = ".") %>%
  rename(region_name = nazwa_miasta)
```

@@

```{r}
selectivv[[1]] %>%
  mutate(month = as.Date(paste('15', month), format='%d %b %Y')) %>%
  gather(apps, vals, -month) %>%
  filter(vals > 0 ) %>%
  mutate(przewoz = apps %in% c("uber_driver", "bolt_driver", 
                               "i_taxi_kierowca_k3", "free_now_for_drivers"),
         przewoz = ifelse(przewoz, "Kierowcy (przewóz osób)", "Dostawcy (przewóz towarów)")) %>%
  mutate(apps = case_when(apps == "bolt_courier" ~"Bolt Courier",
                          apps == "bolt_driver" ~ "Bolt Driver",
                          apps == "free_now_for_drivers" ~ "Free Now",
                          apps == "glover" ~ "Glover",
                          apps == "i_taxi_kierowca_k3" ~ "iTaxi",
                          apps == "takeaway_com_courier" ~ "Takeaway",
                          apps == "uber_driver" ~ "Uber",
                          apps == "wolt_courier_partner" ~ "Wolt"),
         label = if_else(month == max(month), as.character(apps), NA_character_)) %>%
  ggplot(data = ., aes(x = month, y = vals/1000, color = apps, group = apps)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(aes(label = label),
                  nudge_x = -0.25, 
                  direction = "y", 
                  hjust = "left",
                  box.padding = .1,
                  force = .8,
                  na.rm = TRUE,
                  show.legend = FALSE) + 
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  labs(x = "Period", y = 'Number of users (thousands)', color = "App") + 
  facet_wrap(~przewoz, scales = "free_y") +
  scale_x_date(limits = c(as.Date("2018-01-01"), as.Date("2021-09-01"))) + 
  theme(strip.text = element_text(size = 15)) ->  p1

@@

```{r}
selectivv[[2]] %>%
  count(app, country, period, wt = n_unique_users) %>%
  mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                            period == "July - December 2018" ~ "2018.2",
                            period == "January - June 2019" ~ "2019.1",
                            period == "July - December 2019" ~ "2019.2",
                            period == "January - June 2020" ~ "2020.1",
                            period == "July - December 2020" ~ "2020.2"),
         country = factor(country, c("PL", "UA", "other"), c("Poland", "Ukraine", "Other")),
         przewoz = app %in% c("Uber Driver", "Bolt Driver", 
                               "iTaxi Kierowca K3", "FREE NOW for drivers"),
         przewoz = ifelse(przewoz, "Transport", "Delivery"),
         app = case_when(app == "Bolt Courier" ~"Bolt Courier",
                          app == "Bolt Driver" ~ "Bolt Driver",
                          app == "FREE NOW for drivers" ~ "FREE NOW",
                          app == "Glover" ~ "Glover",
                          app == "iTaxi Kierowca K3" ~ "iTaxi",
                          app == "Takeaway.com Courier" ~ "Takeaway",
                          app == "Uber Driver" ~ "Uber",
                          app == "Wolt Courier Partner" ~ "Wolt")) %>%
  ggplot(data = ., aes(x = period, y = n/1000, group = country, color = country)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ przewoz + app) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Country") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2a

@@

```{r}
selectivv[[2]] %>%
  count(app, age, period, wt = n_unique_users) %>%
  mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                            period == "July - December 2018" ~ "2018.2",
                            period == "January - June 2019" ~ "2019.1",
                            period == "July - December 2019" ~ "2019.2",
                            period == "January - June 2020" ~ "2020.1",
                            period == "July - December 2020" ~ "2020.2"),
         #country = factor(country, c("PL", "UA", "other"), c("Poland", "Ukraine", "Other")),
         przewoz = app %in% c("Uber Driver", "Bolt Driver", 
                               "iTaxi Kierowca K3", "FREE NOW for drivers"),
         przewoz = ifelse(przewoz, "Transport", "Delivery"),
         app = case_when(app == "Bolt Courier" ~"Bolt Courier",
                          app == "Bolt Driver" ~ "Bolt Driver",
                          app == "FREE NOW for drivers" ~ "FREE NOW",
                          app == "Glover" ~ "Glover",
                          app == "iTaxi Kierowca K3" ~ "iTaxi",
                          app == "Takeaway.com Courier" ~ "Takeaway",
                          app == "Uber Driver" ~ "Uber",
                          app == "Wolt Courier Partner" ~ "Wolt")) %>%
  ggplot(data = ., aes(x = period, y = n/1000, group = age, color = age)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ przewoz + app) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Age group") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2b

@@

```{r}
selectivv[[2]] %>%
  count(app, gender, period, wt = n_unique_users) %>%
  mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                            period == "July - December 2018" ~ "2018.2",
                            period == "January - June 2019" ~ "2019.1",
                            period == "July - December 2019" ~ "2019.2",
                            period == "January - June 2020" ~ "2020.1",
                            period == "July - December 2020" ~ "2020.2"),
         gender = factor(gender, c("MEN", "WOMEN"), c("Men", "Women")),
         przewoz = app %in% c("Uber Driver", "Bolt Driver", 
                               "iTaxi Kierowca K3", "FREE NOW for drivers"),
         przewoz = ifelse(przewoz, "Transport", "Delivery"),
         app = case_when(app == "Bolt Courier" ~"Bolt Courier",
                          app == "Bolt Driver" ~ "Bolt Driver",
                          app == "FREE NOW for drivers" ~ "FREE NOW",
                          app == "Glover" ~ "Glover",
                          app == "iTaxi Kierowca K3" ~ "iTaxi",
                          app == "Takeaway.com Courier" ~ "Takeaway",
                          app == "Uber Driver" ~ "Uber",
                          app == "Wolt Courier Partner" ~ "Wolt")) %>%
  ggplot(data = ., aes(x = period, y = n/1000, group = gender, color = gender)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ przewoz + app) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Gender") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2c

@@

```{r}
selectivv[[3]]  %>%
    filter(group == "Cities")  %>%
    mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                              period == "July - December 2018" ~ "2018.2",
                              period == "January - June 2019" ~ "2019.1",
                              period == "July - December 2019" ~ "2019.2",
                              period == "January - June 2020" ~ "2020.1",
                              period == "July - December 2020" ~ "2020.2")) %>%
    select(period, region_name, bolt_driver:glover) %>%
    gather(apps, users_cities, bolt_driver:glover) %>%
  left_join(
    bael 
  ) %>%
  filter(users_cities > 0 ) -> cities_bael
```

```{r}
cities_bael %>%
  filter(region_name != "M. Sopot", period == "2020.1") %>%
  filter(!is.na(y))  %>%
  mutate(share = users_cities/y*100,
         share_hi = users_cities / (y - 1.96*rmse)*100,
         share_lo = users_cities / (y + 1.96*rmse)*100,
         region_name = str_remove(region_name, "M. ")) %>%
  mutate(apps = case_when(apps == "bolt_courier" ~"Bolt Courier",
                          apps == "bolt_driver" ~ "Bolt Driver",
                          apps == "free_now_for_drivers" ~ "FREE NOW",
                          apps == "glover" ~ "Glover",
                          apps == "i_taxi_kierowca_k3" ~ "iTaxi",
                          apps == "takeaway_com_courier" ~ "Takeaway",
                          apps == "uber_driver" ~ "Uber",
                          apps == "wolt_courier_partner" ~ "Wolt"),
         apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                               "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
  ggplot(data = ., aes(x = reorder(region_name, -y), y = share, ymin = share_lo, ymax = share_hi)) +
  geom_errorbar() +
  geom_point()  +
  facet_wrap(~apps, ncol = 4, scale = "free_y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5))  +
  #scale_y_continuous(labels = scales::label_percent(accuracy = 2)) +
  labs(x = "City (ordered by the LFS working population size)", 
       y = "Number of users / LFS working population [%]") -> p3

@@

```{r}
selectivv[[4]] %>%
  mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                            period == "July - December 2018" ~ "2018.2",
                            period == "January - June 2019" ~ "2019.1",
                            period == "July - December 2019" ~ "2019.2",
                            period == "January - June 2020" ~ "2020.1", 
                            period == "July - December 2020" ~ "2020.2")) %>%
  gather(times, vals, -period, -app) %>%
  separate(times, c("day", "day2", "hour"), extra = "merge") %>%
  unite("day", c("day", "day2"), sep = "-") %>%
  mutate(stat = ifelse(str_ends(hour, "std"), "Std.Dev.", "Average"),
         vals = vals / 3600,
         hour = str_remove(hour, "\\_std|\\_s"),
         hour = ifelse(str_detect(hour, "18_00_8_00"), "night (18-8)", "day (8-18)"),
         day  = factor(day, c("monday-thursday", "friday-sunday"), c("Mondays-Thursdays", "Fridays-Sundays")),
         app = case_when(app == "Bolt Courier" ~ "Bolt Courier",
                          app == "Bolt Driver" ~ "Bolt Driver",
                          app == "FREE NOW for drivers" ~ "FREE NOW",
                          app == "Glover" ~ "Glover",
                          app == "iTaxi Kierowca K3" ~ "iTaxi",
                          app == "Takeaway.com Courier" ~ "Takeaway",
                          app == "Uber Driver" ~ "Uber",
                          app == "Wolt Courier Partner" ~ "Wolt"),
         app = factor(app, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                             "Bolt Courier", "Takeaway", "Wolt", "Glover")),
         przewoz = app %in% c("Uber", "Bolt Driver", "iTaxi", "FREE NOW"),
         przewoz = ifelse(przewoz, "Transport", "Delivery"))  %>%
  ggplot(data = ., aes(x = period, y = vals, color = app, group = app)) +
  geom_line() +
  geom_point() + 
  facet_nested(day ~ przewoz +  hour + stat)  +
  scale_color_brewer(type = "qual", palette = "Dark2", name = "Apps") +  
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(x = "Half-year", y = "Average time (in hours)") + 
  scale_y_continuous(breaks = seq(0,7,1))  -> p4

@@

```{r}
selectivv[[3]] %>%
    filter(group == "Cities") %>%
    mutate(period = case_when(period == "January - June 2018" ~ "2018.1",
                              period == "July - December 2018" ~ "2018.2",
                              period == "January - June 2019" ~ "2019.1",
                              period == "July - December 2019" ~ "2019.2",
                              period == "January - June 2020" ~ "2020.1",
                              period == "July - December 2020" ~ "2020.2")) %>%
    select(period, region_name, bolt_driver:glover) %>%
    gather(apps, vals, -region_name, -period) %>%
    mutate(apps = case_when(apps == "bolt_courier" ~"Bolt Courier",
                            apps == "bolt_driver" ~ "Bolt Driver",
                            apps == "free_now_for_drivers" ~ "FREE NOW",
                            apps == "glover" ~ "Glover",
                            apps == "i_taxi_kierowca_k3" ~ "iTaxi",
                            apps == "takeaway_com_courier" ~ "Takeaway",
                            apps == "uber_driver" ~ "Uber",
                            apps == "wolt_courier_partner" ~ "Wolt"),
           apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                                 "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
    filter(vals > 0, period == "2020.2")  %>% spread(apps, vals) %>% select(-period) %>% mutate(region_name = str_remove(region_name, "M. ")) %>% xtable(digits = 0) %>% print.xtable(include.rownames = F, NA.string = "--")
```

>>>>

```{r}
selectivv <- list()
selectivv[[1]] <- read_csv(file = "data/gig-table1-monthly-counts-stats.csv")
selectivv[[2]] <- read_csv(file = "data/gig-table2-halfyear-demo-stats.csv")
selectivv[[3]] <- read_csv(file = "data/gig-table3-halfyear-region-stats.csv")
selectivv[[4]] <- read_csv(file = "data/gig-table4-halfyear-activity-stats.csv")
```

BAEL data 
```{r}
bael <- read_csv("data/bael.csv") %>%
  na.omit() %>%
  unite(period, year:yearhalf, sep = ".") %>%
  rename(region_name = nazwa_miasta)
```

@@

```{r}
selectivv[[1]] %>%
  mutate(label = if_else(month == max(month), as.character(apps), NA_character_)) %>%
  ggplot(data = ., aes(x = month, y = number_of_users/1000, color = apps, group = apps)) +
  geom_line(show.legend = FALSE) + 
  geom_text_repel(aes(label = label),
                  nudge_x = -0.25, 
                  direction = "y", 
                  hjust = "left",
                  box.padding = .1,
                  force = .8,
                  na.rm = TRUE,
                  show.legend = FALSE) + 
  scale_color_brewer(type = "qual", palette = "Dark2") + 
  labs(x = "Period", y = 'Number of users (thousands)', color = "App") + 
  facet_wrap(~category, scales = "free_y") +
  scale_x_date(limits = c(as.Date("2018-01-01"), as.Date("2021-09-01"))) + 
  theme(strip.text = element_text(size = 15)) ->  p1

@@

```{r}
selectivv[[2]] %>%
  count(category, apps, country, period, wt = n_unique_users) %>%  
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = country, color = country)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Country") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2a

@@

```{r}
selectivv[[2]] %>%
  count(category, apps, age, period, wt = n_unique_users) %>%
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = age, color = age)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Age group") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2b

@@

```{r}
selectivv[[2]] %>%
  count(category, apps, gender, period, wt = n_unique_users)  %>%
  ggplot(data = ., aes(x = as.character(period), y = n/1000, group = gender, color = gender)) +
  geom_line() +
  geom_point() + 
  scale_color_brewer(type = "qual", palette = "Dark2") +  
  facet_nested(~ category + apps) +
  labs(x = "Half-year", y = "Number of users (in thousands)", color = "Gender") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1),
        legend.position = "bottom") -> p2c

@@

```{r}
selectivv[[3]]  %>%
    filter(group == "Cities")  %>%
    mutate(period = as.character(period)) %>%
    select(category, period, region_name, apps, number_of_users) %>%
    left_join(
      bael %>%
        mutate(region_name = str_remove(region_name, "M. "))
    ) %>%
    filter(number_of_users > 0 ) -> cities_bael
```

```{r}
cities_bael %>%
  filter(region_name != "Sopot", period == "2020.1") %>%
  filter(!is.na(y))  %>%
  mutate(share = number_of_users/y*100,
         share_hi = number_of_users / (y - 1.96*rmse)*100,
         share_lo = number_of_users / (y + 1.96*rmse)*100) %>%
  mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                               "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
  ggplot(data = ., aes(x = reorder(region_name, -y), y = share, ymin = share_lo, ymax = share_hi)) +
  geom_errorbar() +
  geom_point()  +
  facet_wrap(~apps, ncol = 4, scale = "free_y") +
  theme(axis.text.x = element_text(angle = 45, vjust = 1, hjust = 1, size = 5))  +
  #scale_y_continuous(labels = scales::label_percent(accuracy = 2)) +
  labs(x = "City (ordered by the LFS working population size)", 
       y = "Number of users / LFS working population [%]") -> p3

@@

```{r}
selectivv[[4]] %>%
  mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                             "Bolt Courier", "Takeaway", "Wolt", "Glover")),
         day = factor(day, c("Mondays-Thursdays", "Fridays-Sundays")))  %>%
  ggplot(data = ., aes(x = as.character(period), y = activity_time, color = apps, group = apps)) +
  geom_line() +
  geom_point() + 
  facet_nested(day ~ category +  hour + statistic)  +
  scale_color_brewer(type = "qual", palette = "Dark2", name = "Apps") +  
  theme(axis.text.x=element_text(angle = 45, vjust = 1, hjust = 1)) +
  labs(x = "Half-year", y = "Average time (in hours)") + 
  scale_y_continuous(breaks = seq(0,7,1))  -> p4

@@

```{r}
selectivv[[3]] %>%
    filter(group == "Cities") %>%
    mutate(apps = factor(apps, c("Uber", "Bolt Driver", "iTaxi", "FREE NOW",
                                 "Takeaway", "Wolt", "Glover",  "Bolt Courier"))) %>%
    filter(number_of_users > 0, period == "2020.2")  %>% 
    select(period, apps, region_name, number_of_users) %>%
    spread(apps, number_of_users) %>% 
    select(-period) %>% 
    xtable(digits = 0) %>% 
    print.xtable(include.rownames = F, NA.string = "--")
```
